---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# ozroaddeaths [![Travis-CI Build Status](https://travis-ci.org/ropenscilabs/ozroaddeaths.svg?branch=master)](https://travis-ci.org/ropenscilabs/ozroaddeaths) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropenscilabs/ozroaddeaths?branch=master&svg=true)](https://ci.appveyor.com/project/ropenscilabs/ozroaddeaths)


ozroaddeaths is a package that pulls data from the Australian Road 
    Deaths Database, run by the Bureau of Infrastructure, Transport and Regional
    Economics (BITRE). This provides basic details of road transport crash
    fatalities in Australia as reported by the police each month to the State
    and Territory road safety authorities. The details provided in the
    database fall into two groups: 1) the circumstances of the crash, for
    example, date, location, crash type some details regarding the persons
    killed, for example, age, gender and road user group.

## Installation

You can install ozroaddeaths from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ropenscilabs/ozroaddeaths")
```

## Example

This is a basic example which shows you how to solve a common problem:


```{r example-download}
library(ozroaddeaths)
oz_road_fatal_crash <- oz_road_fatal_crash()
oz_road_fatal_fatalities <- oz_road_fatalities()
```
