---
title: "Getting Started"
author: "Adam Gruer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Australian Road Deaths Database provides basic details of road transport crash fatalities in Australia as reported by the police each month to the State and Territory road safety authorities.

Details provided in the database fall into two groups:

* the circumstances of the crash, for example, date, location, crash type

* some details regarding the persons killed, for example, age, gender and road user group.

### When is the data updated?
The fatality data is updated every month.
The heavy vehicle flags (for articulated truck, heavy rigid truck and bus involvement) are only updated each quarter, and are current to within two months. Information for heavy rigid truck involvement in crashes earlier than 2004 is incomplete.
Citations for information derived from this database should include the database name, the webpage, and the date of access.

## Import data from the BITRE website into R
```{r import the data to tibble, message=FALSE, warning=FALSE, cache=TRUE}
library(ozroaddeaths)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggridges)
crashes <- oz_road_fatal_crash()
fatalities <- oz_road_fatalities()
```

## Whate variables are available?

### Crashes

```{r, results='asis'}
knitr::kable(dplyr::as_data_frame(names(crashes)))
knitr::kable(head(crashes))
```

### Fatalities

```{r fatalities first look, results='asis'}
knitr::kable(dplyr::as_data_frame(names(fatalities)))
knitr::kable(head(fatalities))
```

### Plot crashes by year

```{r crash plot by year, width = 10}
crash_plot <- crashes %>%
  ggplot(aes( x =  year, fill = year ))+
  geom_line(stat = "count") +

  theme_minimal() +
  ggtitle("Annual number of fatal car accidents per year")
crash_plot

```



### Plot crashes by year and state

```{r crash plot by year and state, width = 10}

crash_plot +
  scale_y_continuous(trans = "log2") +
  
  facet_wrap(~state) +
   ggtitle("Annual number of fatal car accidents per year and state",
           subtitle = "log2 scale" )


```

### Fatalities by year

```{r fatalities plot by year, width = 10}
fatality_plot <- fatalities %>%
  mutate(
    year = lubridate::year(date_time)) %>%
  ggplot(aes( x =  year, fill = year ))+
  geom_line(stat = "count") +

  theme_minimal() +
  ggtitle("Annual number of road fatalities")
fatality_plot

```

```{r fatalities plot by age, width = 10}
fatality_plot <- fatalities %>%
  filter(gender != "Unknown") %>%
  mutate(
    year = lubridate::year(date_time)) %>%
  ggplot(aes( x =  age, fill = gender )) +

  geom_density() +
  facet_wrap(~gender) +

  theme_minimal() +
  ggtitle("Distribution of road fatalities by age 1989 to 2017")
fatality_plot

```




